---
title: "Assignment1_CarsDataSet"
author: "MadhusudhanMasineni"
date: "1/28/2022"
output: 
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Cars Data Set import and read

The cars dataset downloaded from kaggle and to plot, explore the statistical variables. Due to large dataset minimized to 2000 records of different car models and makers.

https://www.kaggle.com/aishwaryamuthukumar/cars-dataset-audi-bmw-ford-hyundai-skoda-vw

## Read csv file 
```{r carsDetails}
#setwd("C:/Users/madhu/OneDrive/Desktop/MS/RWorking")
#install.packages(readr)
library(readr)
cars <- read.csv("cars_dataset.csv", header = T, sep = ",")

```

# Descriptive Statistics
```{r}
dim(cars)
ls(cars)
head(cars)
str(cars)
```

# Verify the data null

```{r}
any(is.na(cars))
sum(is.na(cars))
colSums(is.na(cars))
```

# plot some models
```{r}
plot(as.factor(cars$model))
plot(as.numeric(cars$price))
```

## plot some statistics of the main variables : quantitative and categorical variables
```{r}

mean(cars$price)
median(cars$mileage)
quantile(cars$price)
sd(cars$price)

max(cars$price)
min(cars$price)

summary(as.factor(cars$Make))
summary(cars$price)

summary(cars$engineSize)
summary(cars$mileage)

```

# car makes and their model corresponding cars price
```{r}
table(cars$model, cars$price, cars$mileage)

```
## include plots

```{r cars, echo=FALSE}
#transform
(summary(cars$mileage))
summary(cars$year)

table(cars$model)
table(cars$fuelType)

```

## Transformation

```{r}
summary(log10(cars$price))
transform(cars$model)

```

# plot 
```{r}
plot(as.factor(cars$model),as.numeric(cars$price))
plot(as.factor(cars$model), as.numeric(cars$mileage))

barplot(table(cars$model, cars$year), main = "year wise making",
        xlab = "No.Of Cars", ylab = "year", las=1, horiz = T, col="gray")

hist(cars$year, freq = FALSE,main = "Cars making over the years", xlab = "years", ylab = "%", col="green")
```

#scatter plot & install ggplot and plot the scatter plot 
```{r}
#install.packages("car")

library(car)
scatterplot(as.factor(cars$model), as.numeric(cars$price))

library(ggplot2)

ggplot(cars) +
 aes(x = model, y = mileage, fill = engineSize, colour = Make) +
 geom_boxplot() +
 scale_fill_gradient() +
 scale_color_hue(direction = 1) +
 theme_minimal()

```
